\section{The ACN-Subsystem} \label{se:acn}

The subsystem {\it acn} provides a new version of the AC-system that
supports evaluation domains
\index{evaluation domains}
and memorization
\index{memorization}
during normalization.
The normalization strategy is controlled by a new global variable NSTR.
NSTR may have three values.
\begin{description}
\item[purenorm:] Standard innermost normalization modulo AC and C is used.
\item[evalsupport:]  Normalization with evaluation support is used
                     interleaving innermost normalizations (modulo AC and C)
                     and evaluations.
\item[memorize:] Innermost normalization modulo AC and C is used supported
                  by reusing memorized (intermediate) results.
                  Memorizing intermediate results uses partial evaluation.
\end{description}
The modules {\tt accsc}, {\tt acde}, {\tt acm}, {\tt acr}, {\tt actto},
{\tt acu}, {\tt acu1} and {\tt acx} are taken from the subsystem {\it ac}.

\subsection{The {\tt achr}-Module}

The module {\tt achr} exports the normalization procedure TNRMLZ.
TNRMLZ normalizes a term according to the setting of the global variable
NSTR as described above.
The module {\tt achr} also provides an innermost normalization procedure
TIREH that reuses precomputed normal forms.
The normal form $t\!\downarrow_{\cal R}$ of a term $t$ is memorized (hashed)
\index{memorization}
if $t$ is ground and the top operator $t(\lambda)$ of $t$  is marked as
``hashable'', i.\,e., if $t(\lambda)$ has a property under indicator DSH.
In the positive case  $t\!\downarrow_{\cal R}$ is stored  as rule
\( t \rightarrow ev(t\!\downarrow_{\cal R} \) where 
$ev(t\!\downarrow_{\cal R})$ denotes the evaluated form of
$t\!\downarrow_{\cal R}$.
The last evaluations step takes only place if there are evaluable function
symbols.

\subsection{The {\tt accmp}-Module}

The module {\tt accmp} extends the {\tt accmp} module of the {\it ac} subsystem
to take advantage of built-in evaluation domains.
In the procedure DTKBC is is possible to determine a term rewriting system
$\cal R$ that is already known to be complete.
If evaluation support is enabled the reduction relation described by this 
$\cal R$ must be contained in the evaluation relation.

\subsection{The Modules {\tt acf} and {\tt accp}}
 
The modules {\tt acf} and {\tt accp} have be adapted form their versions
in the {\it ac} subsystem to  call TNRMLZ instead of TIRDSE.

\subsection{The EV-Rewrite Laboratory}

The main algorithm EV is essentially a modified version of AC allowing to
select among different  normalization procedures.
