\section{The \protect\redux\  Header Files} \label{ap:hd}

The header files provide macros for access functions to substructures of
most data types used in \redux\ and all global \redux-variables. 
The macro techniques used are described in {\em The ReDuX System
Documentation}.


\subsection*{Global Variables: globs.h}
\label{Ap:gv}
The global, safe variables of \redux\ are defined in the header file
{\tt globs.h}.
\verbatiminput{incl/globs.h}

\subsection*{tp.h}

The header file {\tt tp.h} is the standard header file describing the data
structures  {\em term}, {\em operator}, {\em axiom}, {\em data type},
{\em critical pair}, {\em substitution}, {\em assignment list}
and {\em variable dictionary}.
\verbatiminput{incl/tp.h}

\subsection*{parse.h}

The header file {\tt parse.h} contains additional data types used in the
new ReDuX parser.
\verbatiminput{incl/parse.h}

\subsection*{it.h}

The header file {\tt it.h} contains additional data types used in the
interpreter.
\verbatiminput{incl/it.h}

\subsection*{to.h}

The header file {\tt to.h} contains additional data types
needed for the polynomial interpretation orderings.
\verbatiminput{incl/to.h}

\subsection*{ic.h}

The header file {\tt ic.h} contains additional data types used in the
{\it ic}-subsystem only.
It is based on the header file {\tt tp.h}
\verbatiminput{incl/ic.h}

\subsection*{rd.h}

The header file {\tt rd.h} contains additional macros and global variables
used in the {\it rd}-subsystem only.
\verbatiminput{incl/rd.h}

subsection*{ev.h}

The header file {\tt ev.h} contains additional macros and global variables
used in the {\it ev}-subsystem only.
\verbatiminput{incl/ev.h}

\subsection*{evmenu.h}

The header file {\tt evmenu.h} contains additional macros and global variables
used in the main program of the {\it ev}-subsystem only.
It provedes symbolic names for menu choices.
\verbatiminput{incl/evmenu.h}

\subsection*{uc.h}

The header file {\tt uc.h} contains additional macros and global variables
used in the {\it uc}-subsystem only.
\verbatiminput{incl/uc.h}



